\chapter{Appendix Z: Unrelated yet Inspiring Works}

I've been fortunate during my graduate school career to have explored several fields I found fascinating and fun. These offer little insight into Quadrotor Cinematography, but they did influence my thinking and I feel compelled to yot them down somewhere. Here they are:


\section{Computer Science}

\begin{itemize}

\item[Structure and Interpretation of Computer Programs] Abel and Sussman's legendary text on the fundamentals of Computer Programs. This was the first real computer science text I read as part of Berkeley's CS61A course taught by Prof. Brian Harvey. Getting through this was a monumental experience of having my entire view of the world changed on a scale I've never experienced from a book before or since. I always find peace and a mischievous grin in the opening dedication whenever I get disillusioned with my profession. \footnote{This book is dedicated, in respect and admiration, to the spirit that lives in the computer.

``I think that it's extraordinarily important that we in computer science keep fun in computing. When it started out, it was an awful lot of fun. Of course, the paying customers got shafted every now and then, and after a while we began to take their complaints seriously. We began to feel as if we really were responsible for the successful, error-free perfect use of these machines. I don't think we are. I think we're responsible for stretching them, setting them off in new directions, and keeping fun in the house. I hope the field of computer science never loses its sense of fun. Above all, I hope we don't become missionaries. Don't feel as if you're Bible salesmen. The world has too many of those already. What you know about computing other people will learn. Don't feel as if the key to successful computing is only in your hands. What's in your hands, I think and hope, is intelligence: the ability to see the machine as more than when you were first led up to it, that you can make it more.'' -Alan J. Perlis (April 1, 1922-February 7, 1990)
}

\end{itemize}

\section{Graphic Design, Visual Design, Typography}

\begin{itemize}

\item[Thinking with Type]

\item[Tufte's Works] including 

\end{itemize}

\section{Engineering}

\begin{itemize}

\item[The Art of Electronics]

\end{itemize}

\section{History of Silicon Valley and Computer Science}

\begin{itemize}

\item[The New New Thing] Silicon Graphics is particularly dear to me heart. My introduction to SGI machines at a young age set my on the path of computer science, made me the nerd I am today, and arguably defined the direction of my life.

\item[What The Dormouse Said] This tells the story of how a military instrument of the 40s and 50s became a symbol of personal freedom and counterculture thinking. And of course the exploration of a link between Psychedelics and Computer Science is just awesome!

\end{itemize}

\section{Philosophy and Introspection}

\begin{itemize}

\item Way of the Peaceful Warrior

\end{itemize}
