\chapter{Project Ideas}

\section{Deeply-Connected Filtering for an array of range-based position sensors}

\section{Automatic(Semi?) Editing of Aerial Cinematography}

\subsection{Sub Problems}

alignment of data (multimodal feature matching)

finding useful points in footage

matching to a soundtrack

editing UI

\subsection{Cool examples}
Auto-crash
Auto-hangtime

\section{Auto-Transfer of shots from one environment to another}

find correspondences in the geometry between two scenes, then re-purpose a shot from the previous environment to the new one.

\section{Automatic Perching for Autonomous Quads while Filming}

Observation: People enter a different mode of behavior once the quad is fully autonomous and they're using Horus. For example, they'd like to get a live preview halfway-along the spline, then go and iterate on editing in the virtual environment for a bit. Or, they'd like to have the quad wait for a specific moment when they can film an event, but they're not sure when it'll start.

\footnote{ASIDE: What is the NAME of this phenomenon of people going into different modes? Jeff Heer: ``Effect of Latency on user behavior'': ``The Effects of Interactive Latency on Exploratory Visual Analysis'', ``The effect of System Response Time on interactive computer aided problem solving
''}

Hypothesis: We can build a system that finds good places for the quad to perch in the real world. ``Good'' can be defined at someplace where you can a) still see the action you wanted to see and b) can get as quickly as possible to the start of your spline. 

Some ideas: This can tie into Manolis' work on finding affordances. We want to find all the places that affords perching!

\section{Front-mounted gimbals for infinite pan/tilt/roll}

Put a gimbal on the front of an H-quad. Stabilize in 3 axis. Now we can do crazy stuff! The difficulty seems to be yaw stabilization, which is harder now, and balance, because you're fighting gravity. 

Downside: can't keep flying in one orientation of the quad and yaw the camera 360, would have to yaw the quad.

\section{Incorporate GPS availability into path planning algorithm}

Buildings and obstacles throws GPS shadows that prevents us from receiving signal. 